Detecting Stylistic Variation in Pop Production

### 2023-11-21 outputs/paper/paper.qmd
more edits

Our results suggest that we can capture something about Antonoff's music with these variables, and it might, in the end, just be that subjective "ick" that some people feel, and others enjoy, and that we're all just trying to describe in using different methods and tools. With the understanding that what we understand as "data-driven" in recommendation systems is often deeply intertwined with human subjectivity \-- engineers\' understandings of taste and taste-making shape the recommendation systems that they build (@citeSeaver) \-- we contribute to a more informed understanding of how style and tastes can been algorithmically rendered using Spotify's audio features.

@@

Our analysis reveals that the amalgamation of crowdsourced metrics (danceability) and machine-listening variables offers a nuanced approximation of musical style. Interestingly, the subjective nature of crowdsourced metrics, which might appear rudimentary, plays a crucial role in shaping a more intuitive and embodied understanding of music. This synergy between human subjectivity and algorithmic precision not only mirrors the multifaceted nature of musical perception, but also enriches our model's capacity to capture the essence of an artist's style.

While these features offer a quantitative lens to examine music, they cannot fully encapsulate the nuanced and subjective aspects of musical style and creativity. The role of human subjectivity in interpreting and experiencing music remains paramount, as evidenced by the crowdsourced nature of certain features like danceability and valence. This extends to a more nuanced understanding of recommendation systems as a whole. As Nick Seaver anthropological study of music recommendation companies illustrates, engineers \"develop ways of thinking about musical preference, and about software, that try to reconcile them with each other\" (@citeSeaver).

>>>>

Our results suggest that we can capture something about Antonoff's music with these variables, and it might, in the end, just be that subjective "ick" that some people feel, and others enjoy, and that we're all just trying to describe in using different methods and tools. With the understanding that what we understand as "data-driven" in recommendation systems is often deeply intertwined with human subjectivity \-- engineers' understandings of taste and taste-making shape the recommendation systems that they build (@citeSeaver) \-- we contribute to a more informed understanding of how style and tastes can been algorithmically rendered using Spotify's audio features.

@@

Our analysis reveals that the amalgamation of crowdsourced metrics (like danceability) and machine-listening variables offers a nuanced approximation of musical style. This approach not only mirrors the multifaceted nature of auditory perception but also enriches our model's capacity to capture the essence of an artist's style. Interestingly, the subjective nature of crowdsourced metrics, which might initially seem rudimentary, plays a crucial role in this process. They shape a more intuitive and embodied understanding of music, enhancing the overall analysis.

For instance, the work of Martikainen et al. illustrates this point effectively (@podstyle). They combine open-source audio tools with qualitative, crowd-sourced listener data to understand audio-based stylistic variation in podcasts. This synergy between human subjectivity and algorithmic precision underlines the importance of blending these methods for a comprehensive analysis.

However, it's important to note that while these features offer a quantitative lens to examine music, they cannot fully encapsulate the nuanced and subjective aspects of musical style and creativity. The role of human subjectivity is pivotal in interpreting and experiencing music, as evidenced by the crowdsourced nature of certain features like danceability and valence. This aspect becomes even more significant when considering the broader scope of music recommendation systems.

As Nick Seaver's anthropological study of music recommendation companies shows, engineers "develop ways of thinking about musical preference and software, attempting to reconcile them with each other" (@citeSeaver). This extends to a more nuanced understanding of recommendation systems as a whole, highlighting the intricate balance between computational measures and human interpretation in the evolving landscape of music analysis and appreciation.

Future studies should look into reproducing crowdsourced danceability or valence variables to better understand the underlying user preferences that motivate these metrics. In doing so, we can build more robust ways to computationally model more ephemeral and subjective aspects of music such as style.
